---
title: "Forecasting Sword Demand in R"
author: "Damola Mabogunje"
date: "`r Sys.Date()`"
output: html_document
---

```{r Setup CRAN Repo for importing Libraries & Packages}
# Hardcode the US repo for CRAN
r <-getOption("repos");
r["CRAN"] <- "http://cran.us.r-project.org";
options(repos = r);
rm(r);
```

## Introduction

This R Markdown document replicates the work done in chapter 8 of the Data Smart book by John W. Foreman which treats the technique of Time Series Forecasting using Excel. 

Here, we're going to be using R to achieve the same goal, and I hear that it's pretty short and sweet. So hold on to your britches as we import the forecast library and load up the data.

```{r Import forecast library}

install.packages("forecast", dependencies = TRUE);
library(forecast);

```

Now we load and prep the data

```{r Load and Prep Sword Demand Data}

sword <- read.csv("data/SwordDemand.csv");

# Convert to Time Series Data
sword.ts <- ts(sword, frequency = 12, start = c(2010, 1));

# Plot Time Series Data
plot(sword.ts);

```

## Time to Forecast

With the Sword Demand data now in time-series format, we can feed it into the forecast function.

```{r Forecast Away!}

sword.forecast <- forecast(sword.ts);
plot(sword.forecast);

```

## Conclusion

And we're done! In just a few lines we converted a dataset into a Time Series in R, and were able to forecast ahead based on that data. This may be the simplest implementation so far.
